---
title: "Day 01 - Exercise 01 - Intro Exercises"
author: "Victor Calderon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---
```{r echo=FALSE}
rm(list = ls())
```


This are the responses/answers to the problems posed by the exercises in the
`Intro.R` file.

## Exercise 1.1

> Look through the help file for `plot`, then recreate any figure above and set
> the subtitle to `subtitle`.

```{r}
help(plot)
```

Now we can recreate the figure as in `intro.R`:
```{r}
x = rnorm(1000)
y = rnorm(1000)
plot(x, y, sub = 'subtitle')
```


---

## Exercise 1.2

> Find two different expressions to create a 3 x 3 (row x col) matrix with the
> values 2, 4, and 6 in the rows.

---

## Exercise 1.3

> What happens when you collate (`c()`) a list and a vector? List and a list?

---

## Exercise 1.4

> Install and load the `manipulate` package.

---

## Exercise 1.5

> Use the `manipulate` function to interactively vary the `phi` argument to
> `persp` in the above example. Use `persp(x,y,fa,theta=30,phi=phi_slider)` 
> as the first argument (see the `manipulate` help file and examples).

---

## Exercise 1.6

> Generate 10 values from the normal distribution with mean 5 and sd 3 and
> compute their sample mean.

---

## Exercise 1.7

> Use the `replicate` function to repeat item 6. 1000 times.

---

## Exercise 1.8

> Use the `hist` function to plot a histogram of the sample
> means from item 7. Repeat where `N = 50` instead of `10`. Use
> the `add = TRUE` and `col = "red"` arguments to the 'hist'
> function to add the second histogram to the first for
> for comparison.




















