---
title: "Day 01 - Exercise 01 - Intro Exercises"
author: "Victor Calderon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---
```{r echo=FALSE}
rm(list = ls())
```


This are the responses/answers to the problems posed by the exercises in the
`Intro.R` file.

## Exercise 1.1

> Look through the help file for `plot`, then recreate any figure above and set
> the subtitle to `subtitle`.

```{r}
help(plot)
```

Now we can recreate the figure as in `intro.R`:
```{r}
x = rnorm(1000)
y = rnorm(1000)
plot(x, y, sub = 'subtitle')
```


---

## Exercise 1.2

> Find two different expressions to create a 3 x 3 (row x col) matrix with the
> values 2, 4, and 6 in the rows.

The first method:

```{r}
matrix(c(2, 4, 6), nrow = 3, ncol = 3, byrow = TRUE)
```

and the 2nd version:
```{r}
matrix(rep(c(2,4,6), 3), nrow = 3, ncol = 3, byrow = TRUE)
```
and the rest of forms: 
```{r}
matrix(rep(seq(2, 6, 2), 3), nrow = 3, ncol = 3, byrow = TRUE)
matrix(rep(seq(2, 6, 2), c(3, 3, 3)), nrow = 3, ncol = 3)
t(matrix(rep(c(2, 4, 6), 3), 3, 3))
cbind(rep(2, 3), rep(4, 3), rep(6, 3))
rbind(c(2, 4, 6), c(2, 4, 6), c(2, 4, 6))

```

These are just a few ways of determining a matrix of shape (3, 3), with
the elements (2, 4, 6) in each row.

---

## Exercise 1.3

> What happens when you collate (`c()`) a list and a vector? List and a list?

Collating a list and a vector
```{r}
# Defining my list and vector
list1 = list(c(1, 2, 3))
vec1  = c(5, 6)
c(list1, vec1)

```

Now collating a list with a list
```{r}
list2 <- list(c( 1, 2, 3))
list3 <- list(c( 4, 5, 6))
c(list2, list3)
```
or another example:

```{r}
c(list(1,2), c(1,2))
c(list(1,2), list(1,2))
```

---

## Exercise 1.4

> Install and load the `manipulate` package.

Installing the package `manipulate`

with `install.packages("manipulate")`

and loading the package with

`library("manipulate")`

```{r echo=FALSE}
    library("manipulate")
```


---

## Exercise 1.5

> Use the `manipulate` function to interactively vary the `phi` argument to
> `persp` in the above example. Use `persp(x,y,fa,theta=30,phi=phi_slider)` 
> as the first argument (see the `manipulate` help file and examples).

We will vary `phi`:

```{r}
x = seq(-pi, pi, length = 50)
y = x

# Creating function
f = outer(x, y, function(x, y)cos(y)/(1 + x^2))
fa = (f - t(f)) / 2

# Manipulating figure

#manipulate(persp(x,y,fa,theta=30,phi=phi_slider),
#           phi_slider=slider(0, 90))
```


---

## Exercise 1.6

> Generate 10 values from the normal distribution with mean 5 and sd 3 and
> compute their sample mean.

Generating 10 values from a **normal** distribution:

```{r}
mu = 5
sd = 4
random_values = rnorm(10, mean = mu, sd = sd)
head(random_values)

# Mean value
mean_random = mean(random_values)
head(mean_random)
```

---

## Exercise 1.7

> Use the `replicate` function to repeat item 6. 1000 times.

We will now use the `replicate` function:

```{r}
head(replicate(1000, mean(rnorm(10, mean = mu, sd = sd))))
```

---

## Exercise 1.8

> Use the `hist` function to plot a histogram of the sample
> means from item 7. Repeat where `N = 50` instead of `10`. Use
> the `add = TRUE` and `col = "red"` arguments to the 'hist'
> function to add the second histogram to the first for
> for comparison.

```{r}
library(scales)
## 1st histogram
x1 = replicate(1000, mean(rnorm(10, mean = mu, sd = sd)))
hist(x1, col=scales::alpha('skyblue',.8), border = F)

## 2nd histogram
x2 = replicate(1000, mean(rnorm(50, mean = mu, sd = sd)))
hist(x2, add = TRUE, col=scales::alpha('red',.7), border = F)
```



















