---
title: "Day 01 - Exercise 02 - Data Manipulation"
author: "Victor Calderon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r echo=FALSE}
rm(list = ls())
library('reshape2')
```


This are the responses/answers to the problems posed by the exercises in the
`data-manipulation` file.

## Preamble

> Suppose the data below represent body weight measurements 
> for six participants in a weight loss intervention program,
> at three follow-up time points (t0, t1, and t2)

```{r}
dat <- data.frame(id  = c( 1, 2, 3, 4, 5, 6),
                  age = c(47,52,35,28,62,44),
                  sex = factor(c("M","M","F","M","F","F")),
                  wt_t0 = c(278, 340, 239, 290, 244, 220),
                  wt_t1 = c(230, 302, 231, 277, 245, 201),
                  wt_t2 = c(211, 295, 231, 282, 243, 182))

head(dat)
```

---

## Exercise 1

> 1. Convert these data to long format so that there is only one variable
> with weight measurements, and three rows per participant ('id')

Converting to *long format*:

```{r}
data_melt <- melt(dat, id.vars=c('id','age','sex'))
head(data_melt)
```
We can examine the types of variables:
```{r}
unique(data_melt$variable)
```
and showing the lines for *Participant 1*
```{r}
nrows_per_participant <- sum(ifelse(data_melt$id == 1, TRUE, FALSE))
nrows_per_participant
```
This shows that there are **"`r nrows_per_participant`"** rows per participant!

---

## Exercise 2

> Convert back to wide format, averaging across participant sex 
> (resulting data frame should have one row for males and one for females)

Converting back to *wide* format:
```{r}
dcast(data_melt, sex ~ variable, mean)
```

---

## Exercise 3

> Suppose there is a second data frame that contains additional information
> on participants and others, including marital status and other variables.
> Add marital status (only) to the above data (either the wide or long version)
> using the 'merge' command. What type of 'join' is this?

```{r}
demo <- data.frame(id = 1:20,
                   married = sample(c(T,F), 20, replace=TRUE, prob=c(0.5,0.5)),
                   income  = rnorm(20, 75000, 10000),
                   state   = sample(state.abb, 20, replace=TRUE))
```



---

## Exercise 4


> Suppose that a measure of 'irritability' (scale from 0 to 10) was also
> collected at each time point (see data below).
> Use melt and cast to convert this data to long format with three
> rows per participant, and one column for weights, and one column 
> for irritability scores. (hint: after melting, you must add a 
> 'measure' variable to indicate whether the corresponding value 
> is a weight measurement or irritability measurement, and you 
> must also add a time variable (i.e., 0, 1, or 2))

```{r}
dat <- data.frame(id  = c( 1, 2, 3, 4, 5, 6),
                  age = c(47,52,35,28,62,44),
                  sex = factor(c("M","M","F","M","F","F")),
                  wt_t0 = c(278, 340, 239, 290, 244, 220),
                  wt_t1 = c(230, 302, 231, 277, 245, 201),
                  wt_t2 = c(211, 295, 231, 282, 243, 182),
                  irr_t0 = factor(c(0, 1, 1, 0, 0, 0), levels=0:10, ordered=TRUE),
                  irr_t1 = factor(c(5, 3, 2, 1, 5, 7), levels=0:10, ordered=TRUE),
                  irr_t2 = factor(c(4, 3, 3, 1, 4, 6), levels=0:10, ordered=TRUE))
```


---

## Exercise 5

> Repeat the above task using the 'reshape' function.

---


