---
title: "Day 05 - Exercise 01 - Inference and Predictions"
author: "Victor Calderon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Removing previous environment
rm(list = ls())

## Loading packages

```

# Preamble
The exercises here are taken from the `inferences-vs-predictions.R` file in the 
`scripts/day_05` directory.

In order to be able to run the code in here, you need to run the following:

```r
## Installing packages
install.packages('caret')
install.packages('fGarch')
install.packages('boot')

## Loading packages
library('caret')
library('fGarch')
library('boot')
```

These packages are needed in order to run some of the commands.

We can now **load** the necessary libraries:
```{r echo=FALSE, include=FALSE}
library('caret')
library('fGarch')
library('boot')
```

---

# Model assumptions

## Exercise 01

> Redo the simulations above. Instead of heteroscedastic errors, let the 
> errors have a skewed normal distribution (use 'fGarch::rsnorm').

## Exercise 02

> Redo the simulations above. Instead of heteroscedastic errors, let the 
> errors have a normal distribution. But, assume that 'y' is also related
> to a second factor 'z' that is negatively correlated with 'x', and that
> 'z' is unmeasured (can't be used for inference/prediction). 'z' is an
> unmeasured confounder.

---

# Bootstrap

## Exercise 03

> Rerun the simulations above under the unmeasured confounding scenario
> (question 2 in the modeling assumptions lab). Does the bootstrap interval
> have better coverage than the Wald interval?


---

# Sample bias

## Exercise 04

> Redo the simulation above. Does the bootstrap interval have better actual
> coverage than the Wald interval under sample bias?

---
