---
title: "Day 04 - Exercise 02 - Clustering"
author: "Victor Calderon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Removing previous environment
rm(list = ls())

## Loading packages

```

## Preamble
The exercises here are taken from the `clustering.R` file in the 
`scripts/day_04` directory.

In order to be able to run the code in here, you need to run the following:

```r
## Installing packages
install.packages('ISLR')

## Loading packages
library('ISLR')
```

These packages are needed in order to run some of the commands.

We can now **load** the necessary libraries:
```{r echo=TRUE}
library('ISLR')
```

## K-means section

### Exercise 1

> Use the 'NCI60' data from the 'ISLR' package. This is a microarray
> data set with expression measurements on 6830 genes and 64 cancer 
> cell lines. Treat genes as the input variable and perform PCA. Create
> a scree plot to examine the proportion of variability explained. How
> many PCs are needed to explain 80%, 90%, and 95% of the variability 
> in the expression data (note that if p > n, there are at most p PCs)?
> Perform k-means clustering using the PC scores. Use only those 
> PCs needed to achieve 80%, 90%, and 95% of the variability. Use k=4
> to make four clusters. Explore how this PCA dimensionality reduction affects
> how the four clusters of cell lines group by cancer type.

First, you need to load the dataset `NCI60`.
```{r}
library('ISLR')

## Saving `labs` and `data`
nci.labs <- NCI60$labs
nci.data <- NCI60$data
```


---


## Hierarchical clustering

### Exercise 2

> Use hierarchical clustering to repeat the last part of the k-means lab
> question Use only those PCs needed to achieve 80%, 90%, and 95% of the
> variability. Use hclust and cutree to make 4 clusters. Explore how this
> PCA dimensionality reduction affects how the four clusters of cell lines
> group by cancer type.

---

